---
title: "Data import"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
library(haven)
```

##read in some data

Read in the litters dataset
```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

##Take a look at the data

Printing in the console
```{r}
litters_df
```

You can also look at the head or tail(first or last 6 rows)
```{r}
head(litters_df)
tail(litters_df)
```

Gives you a summary of measures 
```{r}
skimr::skim(litters_df)
```

You can also use view(litters_df) but type it in console. It helps you see the whole dataset
clearly in another tab.

##Options to read_csv

```{r}
litters_df = read_csv("./data/FAS_litters.csv", skip = 10, col_names = FALSE)
```
if you want to skip the first x (in this example 10) columns in your dataset becuase
maybe it is blank or not needed.. you can use this. Then you  can add the col_names = false statement so that you can get random headings for the columns.

```{r}
litters_df = read_csv("./data/FAS_litters.csv", na = c("", "NA", ".", "999"))
```

check out '?read_csv ()' for more information and ways to use it


##Other file formats

READ in an excel 

```{r}
mlb_df = read_excel("./data/mlb11.xlsx", range = "A1:F7")
mlb_df
```

Read in a Sas file
```{r}
pulse_df = read_sas("./data/public_pulse_data.sas7bdat")
pulse_df = janitor::clean_names(pulse_df)
pulse_df
```
I used the janitor clean names just to get the names of the columns a bit presentable

Read.csv vs Read_csv

TRY NOT TO USE READ.CSV(WHICH IS BASE R CODE)... THERES A SLIGHT DIFFERENCE B/N THAT AND READ_CSV(WHICH IS NICER WHEN YOU PRINT OUT THE INFORMATION, SOME OTHER THINGS LIKE PARSING AND INCOMPLETE VARIBALE NAMES CAN ALSO BE SEEN)


##Exporting data
Export the mlb subtable. remember that the current version of mlb table we are using
is the smaller one from A1 to F7
```{r}
write_csv(mlb_df, "./data/mlb_subtable.csv")
```

