---
title: "Data Manipulation"
output: github_document
---

```{r}
library(tidyverse)
```

##Load the faslitter data

```{r}
litters_df = read_csv("./data/fas_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

##select

Choose some columns and not others

```{r}
select(litters_df, group, litter_number)
```

Choose a range of columns use colon
```{r}
select(litters_df, group, gd0_weight:pups_born_alive)
```

you can also sepcify somemthing you want to remove
```{r}
select(litters_df, -litter_number)
```


You can also rename specific columns as part of the procedure when selecting them
```{r}
select(litters_df, GROUP = group, LITTERNUMBER = litter_number)
```

You can also just rename instead of selecting
```{r}
rename(litters_df, GROUP = group, LITTERNUMBER = litter_number)
```



Select helpers
```{r}
select(litters_df, starts_with("gd"))
```

Lets say you want to put litter numbers at the first column and also keep everything
```{r}
select(litters_df, litter_number, everything())
```

```{r}
relocate(litters_df, litter_number)
```