---
title: "Data Manipulation"
output: github_document
---

```{r}
library(tidyverse)
```

##Load the faslitter data

```{r}
litters_df = read_csv("./data/fas_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

##select

Choose some columns and not others

```{r}
select(litters_df, group, litter_number)
```

Choose a range of columns use colon
```{r}
select(litters_df, group, gd0_weight:pups_born_alive)
```

you can also sepcify somemthing you want to remove
```{r}
select(litters_df, -litter_number)
```


You can also rename specific columns as part of the procedure when selecting them
```{r}
select(litters_df, GROUP = group, LITTERNUMBER = litter_number)
```

You can also just rename instead of selecting
```{r}
rename(litters_df, GROUP = group, LITTERNUMBER = litter_number)
```



Select helpers
```{r}
select(litters_df, starts_with("gd"))
```

Lets say you want to put litter numbers at the first column and also keep everything
```{r}
select(litters_df, litter_number, everything())
```

```{r}
relocate(litters_df, litter_number)
```

##filter

```{r}
filter(litters_df, gd0_weight >= 22)
```

Note that below he used two equal to signs becuase hes not trying to make that variable 20 hes trying to say if the value is 20 then keep that row.
```{r}
filter(litters_df, gd_of_birth == 20)
```


In the next one the exclamation mark negates so now it will give you all the rows where gdofbirth is not 20
```{r}
filter(litters_df, !(gd_of_birth == 20))
```


In this next one we want to ammend two variables
```{r}
filter(litters_df, gd0_weight >= 22, gd_of_birth == 20)
```


What of the data is not numeric, lets say you want to keep any row with the categorical value "Con7" in it
```{r}
filter(litters_df, group == "Con7")
```


What if you want to keep 2 different categoricals eg Con7 and Mod8
```{r}
filter(litters_df, group %in% c("Con7", "Mod8"))
```



##Mutate

In the example below we will create a new variable called weight gain(gestational weight on day 18 - gestational weight on day 0). We can also in the same statement change the data in the column group to lower case letters)

```{r}
mutate(
  litters_df,
  wt_gain = gd18_weight - gd0_weight)
```

```{r}
mutate(litters_df, 
       wt_gain = gd18_weight - gd0_weight,
       group = str_to_lower(group))
```



